Granular measures of agricultural land-use influence lake nitrogen and phosphorus differently at macroscales

### 2020-04-17 manuscript/manuscript.Rmd
finish inline linking and typo fixes

Our first source for granular agricultural information was the 2010 Cropland Data Layer (CDL, \citealt{usda-nassNationalAgriculturalStatistics2019}), which we used to generate lake watershed and stream buffer estimates of the proportion of land cover due to specific crops such as corn and soybeans. We used the R package cdlTools to access CDL products \citep{chenToolsDownloadWork2018}. We used the 2010 CDL data because this was the first year of high resolution (30 x 30m) CDL data collection that matched our lake nutrient data collection window. Because the CDL contains detailed coverage for dozens of crops, including rare crops with little to no coverage, we re-categorized CDL data based on \citet{larkCroplandExpansionOutpaces2015} to a more limited set of categories (Table A1).

As a more granular representation of watershed agricultural land use, we measured the proportion of land cover in agricultural and “natural” land-uses in 100-m buffers of lake shorelines and streams for each lake watershed. We chose a 100-m buffer width because this width is inclusive of nearly all “riparian buffers” \citep{mayerMetaAnalysisNitrogenRemoval2007}. Here agricultural land-use is defined in the aggregate sense as the proportion of land used for agriculture in a lake’s watershed whereas “natural” land-uses include the sum of all non-agriculture and non-developed CDL classes. We identified the streams associated with each lake using the stream network extraction tool in the nhdR interface \citep{stachelekNhdRToolsWorking2019} to the National Hydrography Network \citep{usgsNationalHydrographyDataset2019}. We compiled soil characteristics for each lake watershed using the Gridded Soil Survey Geographic Database (gSSURGO, \citealt{soil2016gridded}) where we used the python package gssurgo \citep{stachelekGssurgoPythonToolbox2019} to access gSSURGO  products. One such product was wetland potential, defined as the percentage of the soil grid that meets the criteria for hydric soils formed under conditions of saturation, flooding, or ponding long enough to develop anaerobic conditions but not so long as to be classified as a permanent waterbody \citep{soil2016gridded}. Finally, we compiled mean annual nutrient inputs via fertilizer and manure to lake watersheds from 1982-2001 using county level data provided by \citet{ruddy2006county}. We spatially aligned these county level estimates to lake watershed polygons provided by LAGOS-NE \citep{LAGOSgis} using the area-weighted interpolation functions provided by the sf R package \citep{pebesma2018}. All of our data and data processing code are available at \href{https://doi.org/10.5281/zenodo.3754916}{doi:10.5281/zenodo.3754916}.

@@

We restricted the lakes included in the study to those located within the footprint of LAGOS-NE which includes lakes located in 17 Northeastern and Midwest U.S. states \citep{sorannoLAGOSNEMultiscaledGeospatial2017}. We excluded lakes from our analysis if they had a surface area > 400 km2 or a maximum depth > 35 m. These removals resulted in exclusion of approximately 40 lakes which we regarded as outliers that would likely have undue influence on model results because such large and deep lakes are likely to respond differently to enrichment as a result of enhanced stratification. To ensure an adequate comparison between aggregate and granular measures of agriculture, we
further limited lakes in our study to those with at least 10% of the total watershed area devoted to agricultural land-use and those that were sampled at least three times between 2000 and 2010. A total of 928 lakes met these selection criteria (Figure \ref{fig1}). Because we focused on lakes in agricultural watersheds, more than 35  percent of lakes in our study are considered eutrophic to hypereutrophic using chlorophyll a as a diagnostic versus only 15 percent for all lakes from \citet{sorannoLAGOSNEMultiscaledGeospatial2017} located within our study extent (Figure A3). For the regional terms in our models, we used hydrologic regions at the subbasin (i.e. HUC4) level because this level was small enough to give a sense of overall spatial variation but large enough to encompass sufficient numbers of lakes to estimate within region variance.

@@

\noindent where yi is either TN or TP concentration for lake i, and $\beta$ are “global” (i.e. fixed effect) coefficients. This set of global coefficients included estimates for watershed soil characteristics, near-stream land-use, as well as fertilizer and manure inputs for each lake (Xi). Whereas $\beta$ coefficients were estimated as fixed effects, $\gamma$ coefficients and $\alpha$ intercepts were estimated as varying (i.e. random) slopes and intercepts respectively among m hydrologic regions j on watershed land-use (Wij). Region specific intercepts $\alpha$ and $\gamma$ slopes were assumed to come from a multivariate normal distribution (MVN) where $\mu_\alpha$ and $\mu_{\gamma1}$ are their respective grand mean (i.e., population level) estimates. We tested a variety of watershed land-use types for Wij, including both granular and more aggregate measures of agricultural activity. Our only aggregate measure of agricultural activity was agricultural land-use whereas we used several granular measures of agricultural activity. These included both detailed measures of watershed land-use such as corn and soybean cover as well as more detailed measures of nutrient inputs and near-stream land-use (Table \ref{table1}). We included measures of both N and P inputs in both N and P models because of the possibility for stoichiometric interaction. Given our regularization scheme (i.e. our use of “horseshoe priors” described below) there was little reason to exclude P inputs from N models (and vice versa) because unless P inputs are strongly related to lake TN concentration, their coefficient will be forced close to zero.

@@

We fit all models in a Bayesian framework using the brms R package interface to the Stan statistical program \citep{burkner2017brms, stan2017}. We used horseshoe shrinkage priors on all fixed effect coefficients to evaluate variable importance \citep{carvalhoHorseshoeEstimatorSparse2010}. We considered a response variable sensitive to a given predictor if the predictor’s 95% credible interval did not overlap 0. We standardized all predictor variables by subtracting the mean of each variable and dividing by their respective standard deviation so that model coefficients could be compared on roughly the same scale. As a result, the relative sensitivity of a response variable to a particular predictor is related to the relative magnitude of its coefficient estimate. We evaluated model fit of each watershed land-use variable (i.e., each model having one regionally varying coefficient) in two ways. First, we computed a Bayesian $R^2$ following the method of \citet{gelmanRsquaredBayesianRegression2017} and second, we computed differences in expected log predictive density (ELPD) using the leave-one-out cross e validation routines provided by the loo R package and implemented in brms \citep{vehtari2017practical}. The loo package uses leave-one-out cross validation to estimate overall model error by computing O the average error of models iteratively trained on all the data except for a single point. ELPD values have a similar interpretation to information criterion measures such as Akaike information criterion (AIC) or Watanabe-Akaike information criterion (WAIC) except that values are on a different scale \citep{gelmanUnderstandingPredictiveInformation2013}. Typically, models are considered to be different if they are separated by an AIC value of greater than 2 \citep{andersonAvoidingPitfallsWhen2002}, which is equivalent to an ELPD value of -1 \citep{gelmanUnderstandingPredictiveInformation2013}. We report differences in ELPD among models using the notation $\Delta$ELPD. We selected the model with the lowest absolute value ELPD as the “top-ranked” model for detailed reporting and discussion as this signifies the model with the lowest leave one out cross-validation error for N and P respectively.

We used the default settings of brms to generate posterior estimates using four chains of 4,000 iterations each with no thinning and discarding the first 1,000 iterations. We examined model fits to ensure that all models had acceptable convergence of MCMC chains and had approximately normal model residuals. We further tested for spatial correlation among model residuals using the spind R package \citep{carlSpindPackageAccount2018}. All of our code for model fitting and evaluation is available at \href{https://doi.org/10.5281/zenodo.3754916}{doi:10.5281/zenodo.3754916}.

@@

Lake characteristics (e.g., maximum depth and watershed to lake area ratio) along with measures of nutrient transport (e.g., baseflow) and near-stream agriculture were significant predictors in all lake N and P models that we fit. When we  compared among different models for each nutrient individually, we found that those with agricultural watershed O land-use (in the aggregate sense) were top-ranked (i.e., had the lowest absolute value leave-one-out cross validation score) for models of both TN and TP concentrations (Figure 2, Table 2). Although we observed no difference in the specific watershed land-use predictor used in each top-ranked model, we found differences in the extent to which each top-ranked model was substantially different from lower ranked models. For example, in the case of P, all models had nearly identical $R^2$ (0.63) and the difference between the top-ranked model and second ranked model was modest ($\Delta$ELPD = 0.41). For N models, however, agricultural and corn land-use models had higher $R^2$ (0.58) compared to other models and the difference between the top-ranked and second ranked model was more substantial ($\Delta$ELPD = 2.58, Figure 2, Table \ref{table1}).

@@

When we looked more closely at each top-ranked model, we found similarities in the
types of predictors that contributed significantly to the top-ranked N and P models (Figure 3). First, both N and P models included measures of lake characteristics such as maximum depth and watershed to lake area ratio as significant predictors, with the sign of these associated coefficients matching our conventional understanding, in which shallower lakes and lakes with greater hydrologic loads have higher TN and TP concentrations. Second, both models included measures of nutrient transport such as baseflow and precipitation, in which lakes with a “flashier” hydrology (i.e., having lower baseflow) where incoming water is primarily from surface runoff rather than from groundwater, had higher TN and TP concentrations. Finally, both models indicated that high near-stream agriculture (i.e., a high proportion of the area adjacent to the stream network was in agricultural land-use) was associated with lakes having higher TN and TP concentrations. Where N and P models differed was in the effect of soil clay content, in which soils with low clay content were associated with high lake N but had no significant relationship with P.

@@

Although top-ranked N and P models shared some similarities in the type of predictors that contributed significantly to each model, the coefficients of these models differed in magnitude, and thus the top-ranked models varied in their sensitivity to specific predictors (Figure 3). For example, the top-ranked P model was more sensitive to lake characteristics, whereas the top-ranked N model was more sensitive to watershed land-use. Quantitatively, lake TP concentrations were more sensitive to maximum depth ($\beta_{depth}$: median = -0.39, SD = 0.04) compared to lake TN concentrations ($\beta_{depth}$: median = -0.14, SD = 0.04); whereas lake TN concentrations were more sensitive to watershed agriculture land-use ($\beta_{ag}$: median = 0.44, SD = 0.11) compared to lake TP concentrations ($\beta_{ag}$: median = 0.10, SD = 0.08). Finally, although we found that near-stream agriculture was associated with both higher TN and TP concentrations (i.e., a source-effect of near-stream agriculture), there was not a significant difference in the magnitude (i.e., sensitivity) of this coefficient between N ($\beta_{bufferag}$: median = 0.16, SD = 0.06) and P ($\beta_{bufferag}$: median = 0.12, SD = 0.06) models.

No predictors in the nutrient input category appeared to be strongly related to either TN or TP concentrations. One explanation may be that these variables co-varied with watershed and near-stream land-use variables (Figure A5). In an attempt to further investigate this possibility, we fit alternative models excluding all land-use predictors. The results show that, at least in the case of N, removing these predictors e caused model variance to be apportioned from watershed and buffer land-use predictors to N input and P fertilizer predictors (Figure A4). However, this non-land-use N model had a relatively poor fit ($R^2$ = 0.40) compared to the top-ranked model that included land-use as a predictor ($R^2$ = 0.58).

@@

Both TN and TP concentrations were sensitive to measures of watershed land-use as well as near-stream agriculture (Figure 2, Table \ref{table1}). Despite these similarities, we found differences in both the magnitude of these effects and the extent to which we observed regional variation in watershed land-use sensitivity for N and P models (Figure 4). For P, there was little evidence that sensitivity to watershed land-use was regionally variable. More specifically, the credible intervals for the slope of each individual region overlapped the global slope estimate (Figure 4). In contrast, for N we found evidence for regional variation in sensitivity to watershed land-use. For this nutrient, the credible intervals for 2 of the 37 regions did not overlap the global estimate (Figure 4). These regions were found in parts of Iowa, Minnesota, and Illinois and appear to be more sensitive to watershed land-use – i.e., lake N increases at a faster rate per unit increase in agricultural land-use compared to other regions (Figure 5). The median soil clay content of watersheds in these regions was higher than the median across watersheds in all other regions (Figure 6). Furthermore, these two regions had a unique combination of both high soil clay content and extensive tile drainage (Figure S6).

@@

The result showing that lake TN and TP concentrations are sensitive to the spatial configuration of land-use in near-stream buffers is consistent with our expectation and prior research. Specifically, our finding that model coefficients on near-stream agriculture (i.e., agricultural land-use in stream buffers) in all N and P models were significant and positive indicates a nutrient-delivery effect of stream buffer agriculture and suggests that the spatial configuration of agriculture with respect to stream buffers has a detectable influence on both lake N and P at macroscales. This is consistent with prior studies conducted over more limited spatial extents which examined relationships between lake or stream nutrient concentrations and agricultural land-use in stream buffers (Diebel et al. 2007, Baker et al. 2006, Gemesi et al. 2011, Soranno et al. 2015). In contrast to near-stream agriculture, we found that near-stream “natural” land-use was not significant in either N or P models. While we cannot definitively answer why, it may be related to the role that natural buffers play in nutrient cycling for N and P (Alexander et al. 2008, Canham et al. 2012). In the case of N, natural buffers may reduce stream loading by facilitating denitrification, whereas, in the case of P, natural buffers may trap particulate bound material without necessarily removing it (Mayer et al. 2007, Naiman 2010). An alternative possibility is that we are observing a scale effect whereby the delivery effect of near-stream agriculture is spatially consistent whereas the trapping and removal effects of N and P by natural buffers is more spatially variable (Parn et al. 2012). Finally, we may not have observed a protective effect of natural buffers because natural land-use in buffers is too coarse of a proxy for “riparian buffers” composed of forest or herbaceous vegetation (Mayer et al. 2007).

@@

We found that for N models, land-use of specific crops was significant, although it was not found in the top-ranked model. Specifically, we found that watershed land-use in corn production was a significant predictor in the second-ranked N model (Table 2). In the case of P by contrast, neither aggregate nor granular measures of watershed land-use were significant in any models (Figure 2). This can be explained by our finding that all watershed land-use metrics had weak relationships with lake P, especially relative to the strong relationships we observed between P and other factors like lake depth, hydrology, and near-stream agricultural land-use. Here, our finding is consistent with prior studies in stream ecosystems showing that the influence of hydrology exceeds that of agricultural land-use or anthropogenic P inputs (Metson et al. 2017).

@@

None of the nutrient input variables were significant for either N or P models. On the surface, this would seem to contradict the findings of prior research such as that of Bellmore et al. (2018), who found that stream TN concentrations were more strongly controlled by N input predictors relative to measures of either watershed or near-stream land-use. However, our finding has several alternative explanations. First, differences between our lake study and the Bellmore et al. (2018) stream study may simply point to differences in the controls on stream nitrogen concentrations relative to lake nitrogen concentrations (Allan 2004, Canham et al. 2012). Second, our finding that N and P were insensitive to nutrient input variables may be a result of shared variance between nutrient input variables and watershed land-use. This makes sense given that agricultural land-uses and corn land-use in particular are associated with high rates of fertilizer and manure application (Powers 2007). To test this explanation, we formulated N models without a watershed land-use term and observed that model variance was transferred from watershed land-use to the nutrient input terms total N input and P fertilizer (Figure S4). While the positive coefficient on total N input has a straightforward interpretation, the negative coefficient on P fertilizer is unclear. Rather than evidence of a true inverse relationship between P fertilizer and N (or evidence for stoichiometric interaction sensu Paerl et al. 2016), we interpret the negative coefficient on P fertilizer as evidence of either model misspecification (i.e., model fit was poor compared to the model with land-use because N was very sensitive to land-use) or as evidence of multicollinearity among nutrient input variables (Figure S5). Thus, we think that it is likely that the removal of one of the key drivers of N (being land-use) caused model variance to be transferred to nutrient input predictors generally such that the specific highlighting of P fertilizer is a result of noise rather than a true relationship.

@@

An important category of predictors in our models was nutrient transport. For example, we found that baseflow was a significant predictor in all N but especially all P models. This is consistent with prior research at macroscales showing the sensitivity of lake nutrients to this metric (Collins et al. 2017). In contrast to baseflow, we found that watershed soil clay content was a significant predictor in all N models but not in any P models. Furthermore, we found that the coefficient on soil clay content was negative, which was contrary to our expectation, and prior research. It seems to suggest a negative correlation between clay content and N. However, upon closer inspection, we did not find evidence for such a relationship. Instead, we found evidence for a non-linear relationship, which may explain the negative coefficient on clay, whereby clay and TN were positively correlated over most of the range of watershed clay content but were negatively correlated in watersheds with very high soil clay content (> 20%; Figure S7). This non-linear relationship may be an artifact of the specific non-random sampling of our lakes whereby the lakes with extremely high soil clay content watersheds happen to have extremely long water residence times leading to extensive removal of N loads due to denitrification (Groffman et al. 2009).
Overall, granular measures of agriculture were significant in both N and P models. However, the contribution of such measures relative to other non-agricultural predictors was greater for N models (Figure 2, Figure 3). One reason why granular measures of agricultural activity had a greater effect on N models may be that variation in N is less effectively captured solely by lake and watershed characteristics owing in part to the more complex nature of transformations in the nitrogen cycle. This finding is consistent with that of Collins et al. (2017) and Wagner and Schliep (2018) who found that lake depth coefficients were of a much higher magnitude for P relative to N. This is expected since depth strongly controls internal P loading (i.e., recycling), which is a dominant control on lake phosphorus dynamics (Sondergaard et al. 2013).

>>>>

Our first source for granular agricultural information was the 2010 Cropland Data Layer (CDL, \citealt{usda-nassNationalAgriculturalStatistics2019}), which we used to generate lake watershed and stream buffer estimates of the proportion of land cover due to specific crops such as corn and soybeans. We used the R package `cdlTools` to access CDL products \citep{chenToolsDownloadWork2018}. We used the 2010 CDL data because this was the first year of high resolution (30 x 30m) CDL data collection that matched our lake nutrient data collection window. Because the CDL contains detailed coverage for dozens of crops, including rare crops with little to no coverage, we re-categorized CDL data based on \citet{larkCroplandExpansionOutpaces2015} to a more limited set of categories (Table A1).

As a more granular representation of watershed agricultural land use, we measured the proportion of land cover in agricultural and “natural” land-uses in 100-m buffers of lake shorelines and streams for each lake watershed. We chose a 100-m buffer width because this width is inclusive of nearly all “riparian buffers” \citep{mayerMetaAnalysisNitrogenRemoval2007}. Here agricultural land-use is defined in the aggregate sense as the proportion of land used for agriculture in a lake’s watershed whereas “natural” land-uses include the sum of all non-agriculture and non-developed CDL classes. We identified the streams associated with each lake using the stream network extraction tool in the `nhdR` interface \citep{stachelekNhdRToolsWorking2019} to the National Hydrography Network \citep{usgsNationalHydrographyDataset2019}. We compiled soil characteristics for each lake watershed using the Gridded Soil Survey Geographic Database (gSSURGO, \citealt{soil2016gridded}) where we used the python package `gssurgo` \citep{stachelekGssurgoPythonToolbox2019} to access gSSURGO  products. One such product was wetland potential, defined as the percentage of the soil grid that meets the criteria for hydric soils formed under conditions of saturation, flooding, or ponding long enough to develop anaerobic conditions but not so long as to be classified as a permanent waterbody \citep{soil2016gridded}. Finally, we compiled mean annual nutrient inputs via fertilizer and manure to lake watersheds from 1982-2001 using county level data provided by \citet{ruddy2006county}. We spatially aligned these county level estimates to lake watershed polygons provided by LAGOS-NE \citep{LAGOSgis} using the area-weighted interpolation functions provided by the `sf` R package \citep{pebesma2018}. All of our data and data processing code are available at \href{https://doi.org/10.5281/zenodo.3754916}{doi:10.5281/zenodo.3754916}.

@@

We restricted the lakes included in the study to those located within the footprint of LAGOS-NE which includes lakes located in 17 Northeastern and Midwest U.S. states \citep{sorannoLAGOSNEMultiscaledGeospatial2017}. We excluded lakes from our analysis if they had a surface area > 400 $km^2$ or a maximum depth > 35 m. These removals resulted in exclusion of approximately 40 lakes which we regarded as outliers that would likely have undue influence on model results because such large and deep lakes are likely to respond differently to enrichment as a result of enhanced stratification. To ensure an adequate comparison between aggregate and granular measures of agriculture, we
further limited lakes in our study to those with at least 10% of the total watershed area devoted to agricultural land-use and those that were sampled at least three times between 2000 and 2010. A total of 928 lakes met these selection criteria (Figure \ref{fig1}). Because we focused on lakes in agricultural watersheds, more than 35  percent of lakes in our study are considered eutrophic to hypereutrophic using chlorophyll a as a diagnostic versus only 15 percent for all lakes from \citet{sorannoLAGOSNEMultiscaledGeospatial2017} located within our study extent (Figure A3). For the regional terms in our models, we used hydrologic regions at the subbasin (i.e. HUC4) level because this level was small enough to give a sense of overall spatial variation but large enough to encompass sufficient numbers of lakes to estimate within region variance.

@@

\noindent where yi is either TN or TP concentration for lake i, and $\beta$ are “global” (i.e. fixed effect) coefficients. This set of global coefficients included estimates for watershed soil characteristics, near-stream land-use, as well as fertilizer and manure inputs for each lake ($X_i$). Whereas $\beta$ coefficients were estimated as fixed effects, $\gamma$ coefficients and $\alpha$ intercepts were estimated as varying (i.e. random) slopes and intercepts respectively among m hydrologic regions j on watershed land-use ($W_{ij}$). Region specific intercepts $\alpha$ and $\gamma$ slopes were assumed to come from a multivariate normal distribution (MVN) where $\mu_\alpha$ and $\mu_{\gamma1}$ are their respective grand mean (i.e., population level) estimates. We tested a variety of watershed land-use types for $W_{ij}$, including both granular and more aggregate measures of agricultural activity. Our only aggregate measure of agricultural activity was agricultural land-use whereas we used several granular measures of agricultural activity. These included both detailed measures of watershed land-use such as corn and soybean cover as well as more detailed measures of nutrient inputs and near-stream land-use (Table \ref{table1}). We included measures of both N and P inputs in both N and P models because of the possibility for stoichiometric interaction. Given our regularization scheme (i.e. our use of “horseshoe priors” described below) there was little reason to exclude P inputs from N models (and vice versa) because unless P inputs are strongly related to lake TN concentration, their coefficient will be forced close to zero.

@@

We fit all models in a Bayesian framework using the `brms` R package interface to the Stan statistical program \citep{burkner2017brms, stan2017}. We used horseshoe shrinkage priors on all fixed effect coefficients to evaluate variable importance \citep{carvalhoHorseshoeEstimatorSparse2010}. We considered a response variable sensitive to a given predictor if the predictor’s 95% credible interval did not overlap 0. We standardized all predictor variables by subtracting the mean of each variable and dividing by their respective standard deviation so that model coefficients could be compared on roughly the same scale. As a result, the relative sensitivity of a response variable to a particular predictor is related to the relative magnitude of its coefficient estimate. We evaluated model fit of each watershed land-use variable (i.e., each model having one regionally varying coefficient) in two ways. First, we computed a Bayesian $R^2$ following the method of \citet{gelmanRsquaredBayesianRegression2017} and second, we computed differences in expected log predictive density (ELPD) using the leave-one-out cross e validation routines provided by the `loo` R package and implemented in brms \citep{vehtari2017practical}. The `loo` package uses leave-one-out cross validation to estimate overall model error by computing the average error of models iteratively trained on all the data except for a single point. ELPD values have a similar interpretation to information criterion measures such as Akaike information criterion (AIC) or Watanabe-Akaike information criterion (WAIC) except that values are on a different scale \citep{gelmanUnderstandingPredictiveInformation2013}. Typically, models are considered to be different if they are separated by an AIC value of greater than 2 \citep{andersonAvoidingPitfallsWhen2002}, which is equivalent to an ELPD value of -1 \citep{gelmanUnderstandingPredictiveInformation2013}. We report differences in ELPD among models using the notation $\Delta$ELPD. We selected the model with the lowest absolute value ELPD as the “top-ranked” model for detailed reporting and discussion as this signifies the model with the lowest leave one out cross-validation error for N and P respectively.

We used the default settings of brms to generate posterior estimates using four chains of 4,000 iterations each with no thinning and discarding the first 1,000 iterations. We examined model fits to ensure that all models had acceptable convergence of MCMC chains and had approximately normal model residuals. We further tested for spatial correlation among model residuals using the `spind` R package \citep{carlSpindPackageAccount2018}. All of our code for model fitting and evaluation is available at \href{https://doi.org/10.5281/zenodo.3754916}{doi:10.5281/zenodo.3754916}.

@@

Lake characteristics (e.g., maximum depth and watershed to lake area ratio) along with measures of nutrient transport (e.g., baseflow) and near-stream agriculture were significant predictors in all lake N and P models that we fit. When we  compared among different models for each nutrient individually, we found that those with agricultural watershed land-use (in the aggregate sense) were top-ranked (i.e., had the lowest absolute value leave-one-out cross validation score) for models of both TN and TP concentrations (Figure \ref{fig2, Table \ref{table2}). Although we observed no difference in the specific watershed land-use predictor used in each top-ranked model, we found differences in the extent to which each top-ranked model was substantially different from lower ranked models. For example, in the case of P, all models had nearly identical $R^2$ (0.63) and the difference between the top-ranked model and second ranked model was modest ($\Delta$ELPD = 0.41). For N models, however, agricultural and corn land-use models had higher $R^2$ (0.58) compared to other models and the difference between the top-ranked and second ranked model was more substantial ($\Delta$ELPD = 2.58, Figure \ref{fig2}, Table \ref{table1}).

@@

When we looked more closely at each top-ranked model, we found similarities in the
types of predictors that contributed significantly to the top-ranked N and P models (Figure \ref{fig3}). First, both N and P models included measures of lake characteristics such as maximum depth and watershed to lake area ratio as significant predictors, with the sign of these associated coefficients matching our conventional understanding, in which shallower lakes and lakes with greater hydrologic loads have higher TN and TP concentrations. Second, both models included measures of nutrient transport such as baseflow and precipitation, in which lakes with a “flashier” hydrology (i.e., having lower baseflow) where incoming water is primarily from surface runoff rather than from groundwater, had higher TN and TP concentrations. Finally, both models indicated that high near-stream agriculture (i.e., a high proportion of the area adjacent to the stream network was in agricultural land-use) was associated with lakes having higher TN and TP concentrations. Where N and P models differed was in the effect of soil clay content, in which soils with low clay content were associated with high lake N but had no significant relationship with P.

@@

Although top-ranked N and P models shared some similarities in the type of predictors that contributed significantly to each model, the coefficients of these models differed in magnitude, and thus the top-ranked models varied in their sensitivity to specific predictors (Figure \ref{fig3}). For example, the top-ranked P model was more sensitive to lake characteristics, whereas the top-ranked N model was more sensitive to watershed land-use. Quantitatively, lake TP concentrations were more sensitive to maximum depth ($\beta_{depth}$: median = -0.39, SD = 0.04) compared to lake TN concentrations ($\beta_{depth}$: median = -0.14, SD = 0.04); whereas lake TN concentrations were more sensitive to watershed agriculture land-use ($\beta_{ag}$: median = 0.44, SD = 0.11) compared to lake TP concentrations ($\beta_{ag}$: median = 0.10, SD = 0.08). Finally, although we found that near-stream agriculture was associated with both higher TN and TP concentrations (i.e., a source-effect of near-stream agriculture), there was not a significant difference in the magnitude (i.e., sensitivity) of this coefficient between N ($\beta_{bufferag}$: median = 0.16, SD = 0.06) and P ($\beta_{bufferag}$: median = 0.12, SD = 0.06) models.

No predictors in the nutrient input category appeared to be strongly related to either TN or TP concentrations. One explanation may be that these variables co-varied with watershed and near-stream land-use variables (Figure A5). In an attempt to further investigate this possibility, we fit alternative models excluding all land-use predictors. The results show that, at least in the case of N, removing these predictors caused model variance to be apportioned from watershed and buffer land-use predictors to N input and P fertilizer predictors (Figure A4). However, this non-land-use N model had a relatively poor fit ($R^2$ = 0.40) compared to the top-ranked model that included land-use as a predictor ($R^2$ = 0.58).

@@

Both TN and TP concentrations were sensitive to measures of watershed land-use as well as near-stream agriculture (Figure \ref{fig2}, Table \ref{table1}). Despite these similarities, we found differences in both the magnitude of these effects and the extent to which we observed regional variation in watershed land-use sensitivity for N and P models (Figure \ref{fig4}). For P, there was little evidence that sensitivity to watershed land-use was regionally variable. More specifically, the credible intervals for the slope of each individual region overlapped the global slope estimate (Figure \ref{fig4}). In contrast, for N we found evidence for regional variation in sensitivity to watershed land-use. For this nutrient, the credible intervals for 2 of the 37 regions did not overlap the global estimate (Figure \ref{fig4}). These regions were found in parts of Iowa, Minnesota, and Illinois and appear to be more sensitive to watershed land-use – i.e., lake N increases at a faster rate per unit increase in agricultural land-use compared to other regions (Figure \ref{fig5}). The median soil clay content of watersheds in these regions was higher than the median across watersheds in all other regions (Figure \ref{fig6}). Furthermore, these two regions had a unique combination of both high soil clay content and extensive tile drainage (Figure S6).

@@

The result showing that lake TN and TP concentrations are sensitive to the spatial configuration of land-use in near-stream buffers is consistent with our expectation and prior research. Specifically, our finding that model coefficients on near-stream agriculture (i.e., agricultural land-use in stream buffers) in all N and P models were significant and positive indicates a nutrient-delivery effect of stream buffer agriculture and suggests that the spatial configuration of agriculture with respect to stream buffers has a detectable influence on both lake N and P at macroscales. This is consistent with prior studies conducted over more limited spatial extents which examined relationships between lake or stream nutrient concentrations and agricultural land-use in stream buffers \citep{diebelLandscapePlanningAgricultural2009, bakerImprovedMethodsQuantifying2006,gemesiEffectsWatershedConfiguration2011, sorannoEffectsLandUse2015}. In contrast to near-stream agriculture, we found that near-stream “natural” land-use was not significant in either N or P models. While we cannot definitively answer why, it may be related to the role that natural buffers play in nutrient cycling for N and P \citep{alexanderDifferencesPhosphorusNitrogen2008,canhamNitrogenDepositionLake2012}. In the case of N, natural buffers may reduce stream loading by facilitating denitrification, whereas, in the case of P, natural buffers may trap particulate bound material without necessarily removing it \citep{mayerMetaAnalysisNitrogenRemoval2007,naiman2010riparia}. An alternative possibility is that we are observing a scale effect whereby the delivery effect of near-stream agriculture is spatially consistent whereas the trapping and removal effects of N and P by natural buffers is more spatially variable \citep{parnIndicatorsNutrientsTransport2012}. Finally, we may not have observed a protective effect of natural buffers because natural land-use in buffers is too coarse of a proxy for “riparian buffers” composed of forest or herbaceous vegetation \citep{mayerMetaAnalysisNitrogenRemoval2007}.

@@

We found that for N models, land-use of specific crops was significant, although it was not found in the top-ranked model. Specifically, we found that watershed land-use in corn production was a significant predictor in the second-ranked N model (Table \ref{table2}). In the case of P by contrast, neither aggregate nor granular measures of watershed land-use were significant in any models (Figure \ref{fig2}). This can be explained by our finding that all watershed land-use metrics had weak relationships with lake P, especially relative to the strong relationships we observed between P and other factors like lake depth, hydrology, and near-stream agricultural land-use. Here, our finding is consistent with prior studies in stream ecosystems showing that the influence of hydrology exceeds that of agricultural land-use or anthropogenic P inputs \citep{metsonLinkingTerrestrialPhosphorus2017}.

@@

None of the nutrient input variables were significant for either N or P models. On the surface, this would seem to contradict the findings of prior research such as that of \citet{bellmoreNitrogenInputsDrive2018}, who found that stream TN concentrations were more strongly controlled by N input predictors relative to measures of either watershed or near-stream land-use. However, our finding has several alternative explanations. First, differences between our lake study and the \citet{bellmoreNitrogenInputsDrive2018} stream study may simply point to differences in the controls on stream nitrogen concentrations relative to lake nitrogen concentrations \citep{allanLandscapesRiverscapesInfluence2004,canhamNitrogenDepositionLake2012}. Second, our finding that N and P were insensitive to nutrient input variables may be a result of shared variance between nutrient input variables and watershed land-use. This makes sense given that agricultural land-uses and corn land-use in particular are associated with high rates of fertilizer and manure application \citep{powersNutrientLoadsSurface2007}. To test this explanation, we formulated N models without a watershed land-use term and observed that model variance was transferred from watershed land-use to the nutrient input terms total N input and P fertilizer (Figure S4). While the positive coefficient on total N input has a straightforward interpretation, the negative coefficient on P fertilizer is unclear. Rather than evidence of a true inverse relationship between P fertilizer and N (or evidence for stoichiometric interaction sensu \citealt{paerlItTakesTwo2016}, we interpret the negative coefficient on P fertilizer as evidence of either model misspecification (i.e., model fit was poor compared to the model with land-use because N was very sensitive to land-use) or as evidence of multicollinearity among nutrient input variables (Figure S5). Thus, we think that it is likely that the removal of one of the key drivers of N (being land-use) caused model variance to be transferred to nutrient input predictors generally such that the specific highlighting of P fertilizer is a result of noise rather than a true relationship.

@@

An important category of predictors in our models was nutrient transport. For example, we found that baseflow was a significant predictor in all N but especially all P models. This is consistent with prior research at macroscales showing the sensitivity of lake nutrients to this metric \citep{collinsLakeNutrientStoichiometry2017}. In contrast to baseflow, we found that watershed soil clay content was a significant predictor in all N models but not in any P models. Furthermore, we found that the coefficient on soil clay content was negative, which was contrary to our expectation, and prior research. It seems to suggest a negative correlation between clay content and N. However, upon closer inspection, we did not find evidence for such a relationship. Instead, we found evidence for a non-linear relationship, which may explain the negative coefficient on clay, whereby clay and TN were positively correlated over most of the range of watershed clay content but were negatively correlated in watersheds with very high soil clay content (> 20%; Figure S7). This non-linear relationship may be an artifact of the specific non-random sampling of our lakes whereby the lakes with extremely high soil clay content watersheds happen to have extremely long water residence times leading to extensive removal of N loads due to denitrification \citep{groffmanChallengesIncorporatingSpatially2009}.
Overall, granular measures of agriculture were significant in both N and P models. However, the contribution of such measures relative to other non-agricultural predictors was greater for N models (Figure \ref{fig2}, Figure \ref{fig3}). One reason why granular measures of agricultural activity had a greater effect on N models may be that variation in N is less effectively captured solely by lake and watershed characteristics owing in part to the more complex nature of transformations in the nitrogen cycle. This finding is consistent with that of \citet{collinsLakeNutrientStoichiometry2017} and \citet{wagnerCombiningNutrientProductivity2018} who found that lake depth coefficients were of a much higher magnitude for P relative to N. This is expected since depth strongly controls internal P loading (i.e., recycling), which is a dominant control on lake phosphorus dynamics \citep{sondergaardPersistentInternalPhosphorus2013}.
